Drivers of spatial synchrony in giant kelp populations

### 2021-06-05 paper/SupMat.Rnw
adding new decomp

%***Tentative author order
\author{Lawrence W. Sheppard}
\author{Daniel C. Reuman}
\author{Tom Bell}
\author{Jonathan Walter}
\author{Kyle Cavanaugh}
\author{Max Castorani}

@@

@@

\begin{equation}
\bar r_1-\bar r_2 = \Delta^0+\Delta^E+\Delta^C+\Delta^{(EC)}+\Delta^{(E\#C)}, \label(IGR)
\end{equation}

@@

\begin{align}
\Delta^E &= \bar \epsilon_1^E - \bar \epsilon_2^E\\
&=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) - \epsilon_2^0]\\
&=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - \EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) \\
\end{align}

\begin{align}
\Delta^C &= \bar \epsilon_1^C - \bar \epsilon_2^C\\
&=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) - \epsilon_2^0]\\
&=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - \EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) \\
\end{align}

@@

By substituting these expansions into \ref(IGR), we get,

@@

Fixed number, base growth rate

\begin{equation}
\epsilon_i^0=\ln(1-\delta + \delta \frac{\bar B_i}{\bar B_j}) \label{null_i}
\end{equation}

\begin{equation}
\epsilon_i^E=\ln(1-\delta + \delta \frac{B_i}{\bar B_j}) -  \epsilon_i^0 \label{Evar}
\end{equation}

contribution of variance in E

\begin{equation}
\epsilon_i^C=\ln(1-\delta + \delta \frac{\bar B_i}{B_j}) -  \epsilon_i^0 \label{Cvar}
\end{equation}

contribution of variance in C

interaction effect of both E and C varying

\begin{equation}
\bar r_i=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))\label{rbari}
\end{equation}

\begin{equation}
\bar r_i=\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C + \bar \epsilon_i^{EC} \label{rbari_decomp}
\end{equation}

But some of $\bar \epsilon_i^{EC}$ comes from covariance between E and C and some comes from both varying.

\begin{equation}
\bar \epsilon_i^{(E\#C)}=\EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \label{ECvar_sharp}
\end{equation}

\begin{align}
\bar \epsilon_i^{(EC)}&=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) +[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]
\label{r_iSE}
\end{align}

checks out

Need to find $\Delta_1^0, \Delta_1^E, \Delta_1^C, \Delta_1^{(EC)}, \Delta_1^{(E\#C)}$ (where species 1 is the invader and species 2 is the resident)

>>>>

%***Tentative author order
\author{Lawrence W. Sheppard}
\author{Daniel C. Reuman}

@@

\section{E and C definitions}

The lottery model is given by

\begin{equation}
N_i(t+1)=(1-\delta)N(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}.  
\end{equation}

REF also defines $E_i = B_i$ and $C_1=C_2=\frac{B_1 N_1 + B_2 N_2}{\delta N}$. Then in the invader-resident scenario which defines the invader to be essentially zero and $N_2 = N$, $C_1 = C_2 = \frac{B_2}{\delta}$. The growth rate equation relates the two variables of $E$ and $C$ by

\begin{equation}
\bar r_i=\EX(\ln(1-\delta + \frac{E_i}{C_i})).\label{rbarEC}
\end{equation}

\noindent By substituting the definitions,

\begin{align}
\bar r_i &=\EX(\ln(1-\delta + \frac{B_i}{B_2/ \delta})) \\
&= \EX(\ln(1-\delta +\delta \frac{B_i}{B_2})) 
\end{align}

The null growth rate of species $i$ is computed by setting both E and C at their means.

\begin{align}
\epsilon_i^0 &=\ln(1-\delta + \frac{\bar E_i}{\bar C_i}) \\
&=\ln(1-\delta + \delta \frac{\bar B_i}{\bar B_2}) 
\end{align}

The contribution of variance E in to the IGR is computed by setting C to its mean and letting E vary subtracted by the null.

\begin{align}
\epsilon_i^E &=\ln(1-\delta + \frac{E_i}{\bar C_i}) -  \epsilon_i^0 \\
&=\ln(1-\delta + \delta \frac{B_i}{\bar B_2}) -  \epsilon_i^0 
\end{align}

Similarly, the contribution of variance C in to the IGR is computed by setting E to its mean and letting C vary subtracted by the null.

\begin{align}
\epsilon_i^C &=\ln(1-\delta + \frac{\bar E_i}{C_i}) -  \epsilon_i^0 \\
 &=\ln(1-\delta + \delta \frac{\bar B_i}{ B_2}) -  \epsilon_i^0 
\end{align}

The contribution of both E and C varying to the IGR is computed by letting both variables vary subtracted by the null and the contributions of only one variable varying.

So,

\begin{equation}
\bar r_i=\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C + \bar \epsilon_i^{EC}. \label{rbari_decomp}
\end{equation}

But some of $\bar \epsilon_i^{EC}$ comes from covariance between E and C and some comes from both varying independently.

\begin{align}
\bar \epsilon_i^{(E\#C)} &=\EX(\ln(1-\delta + \frac{E_i}{C_i^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] \\
 &=\EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C] 
\end{align}

\begin{align}
\bar \epsilon_i^{(EC)} &=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
&= \EX(\ln(1-\delta + \frac{E_i}{C_i}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
&- [\EX(\ln(1-\delta + \frac{E_i}{C_i^{\#}})) -[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]] \\
&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_2}))-\epsilon_i^0 -\bar \epsilon_i^E - \bar \epsilon_i^C\\
&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}})) +\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C \\
&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_2}))-\EX(\ln(1-\delta + \delta \frac{B_i}{B_2^{\#}}))\\
\end{align}

@@

\begin{equation}
\bar r_1-\bar r_2 = \Delta^0+\Delta^E+\Delta^C+\Delta^{(EC)}+\Delta^{(E\#C)}, \label{IGR}
\end{equation}

@@

\begin{align}
\Delta^E &= \bar \epsilon_1^E - \bar \epsilon_2^E\\
&=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) - \epsilon_2^0]\\
&=\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0 - \EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) \\
\end{align}

\begin{align}
\Delta^C &= \bar \epsilon_1^C - \bar \epsilon_2^C\\
&=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) - \epsilon_2^0]\\
&=\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0 - \EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) \\
\end{align}

@@

By substituting these expansions into \ref{IGR}, we get,

@@

\section{new decomposition term}

A new decomposition term was proposed to quantify the effect of asymmetric associations to coexistence. Since asymmetry only differs in the storage effect term, $\bar \epsilon^{(EC)}$, the contribution of asymmetry associations between the species will be a further decomposition.

\begin{equation}
\bar \epsilon^{(\dot {EC})} = \bar \epsilon^{(EC)} - \bar \epsilon^{(\ddot {EC})}
\end{equation}

The contribution of asymmetry to the storage effect is the observed storage effect subtracted by a null storage effect if the association between species was completely symmetrical. 
The descrition of $\bar \epsilon^{(\ddot {EC})}$ is given in section 3 and equations 9 and 21-23.
